---
title:  "Final Project Basic Outline"
author: "Dave Friedman"
output: html_notebook
---

```{r}
library(scales)
library(textdata)
library(tidytext)
library(tidyverse)
```

## Research Question

State your general specific question and why it is interesting.

## Data Processing

* [Aristotle's Physics](https://classics.mit.edu/Aristotle/physics.html)
* [Aristotle's Metaphysics](https://classics.mit.edu/Aristotle/metaphysics.html)

### Data Intake

#### Plato's _Timaeus_

```{r}
timaeusE <- base::readLines(con = 'timaeusE.txt')
```

```{r}
base::length(timaeusE)
```

```{r}
stringr::str_sub(timaeusE, 1, 25)[1:25]
```

#### Aristotle's _Physics_

```{r}
physicsE <- base::readLines(con = 'physicsE.txt')
```

```{r}
base::length(physicsE)
```

```{r}
stringr::str_sub(physicsE, 1, 25)[1:25]
```

#### Aristotle's _Metaphysics_

```{r}
metaphysicsE <- base::readLines(con = 'metaphysicsE.txt')
```

```{r}
base::length(metaphysicsE)
```

```{r}
stringr::str_sub(metaphysicsE, 1, 25)[1:25]
```

### Data Wrangling

#### Plato's _Timaeus_

```{r}
timaeusE  <- timaeusE[! timaeusE %in% c('', '\t')]
```

```{r}
base::length(timaeusE[! timaeusE %in% c('', '\t')])
```

```{r}
stringr::str_length(timaeusE)
```

```{r}
base::sum(stringr::str_length(timaeusE))
```

```{r}
timaeusE_df <- tibble::tibble(section_num = 1:base::length(timaeusE), section = timaeusE)
```

```{r}
tidy_timaeusE <-
  timaeusE_df %>%
    tidytext::unnest_tokens(output = word, input = section, to_lower = TRUE) %>%
    dplyr::anti_join(tidytext::stop_words, by = 'word') %>%
    dplyr::anti_join(tibble::tibble(word = base::c('cf', 'ff', 'i.e', 'e.g', '1', '2', '3')), by = 'word')
```
  
```{r}
tidy_timaeusE %>%
  dplyr::count(word, sort = TRUE) %>%
  utils::head(10)
```

```{r}
tidy_timaeusE %>%
  dplyr::count(word, sort = TRUE) %>%
  dplyr::filter(n > 30) %>%
  dplyr::mutate(word = stats::reorder(word, n)) %>%
  ggplot2::ggplot(ggplot2::aes(n, word)) +
    geom_col() +
    labs(y = NULL)
```

#### Aristotle's _Physics_

```{r}
physicsE  <- physicsE[! physicsE %in% c('', '\t')]
```

```{r}
base::length(physicsE[! physicsE %in% c('', '\t')])
```

```{r}
stringr::str_length(physicsE)
```

```{r}
base::sum(stringr::str_length(physicsE))
```

```{r}
physicsE_df <- tibble::tibble(section_num = 1:base::length(physicsE), section = physicsE)
```

```{r}
tidy_physicsE <-
  physicsE_df %>%
    tidytext::unnest_tokens(output = word, input = section, to_lower = TRUE) %>%
    dplyr::anti_join(tidytext::stop_words, by = 'word') %>%
    dplyr::anti_join(tibble::tibble(word = base::c('cf', 'ff', 'i.e', 'e.g', '1', '2', '3')), by = 'word')
```
  
```{r}
tidy_physicsE %>%
  dplyr::count(word, sort = TRUE) %>%
  utils::head(10)
```

```{r}
tidy_physicsE %>%
  dplyr::count(word, sort = TRUE) %>%
  dplyr::filter(n > 45) %>%
  dplyr::mutate(word = stats::reorder(word, n)) %>%
  ggplot2::ggplot(ggplot2::aes(n, word)) +
    geom_col() +
    labs(y = NULL)
```

#### Aristotle's _Metaphysics_

```{r}
metaphysicsE  <- metaphysicsE[! metaphysicsE %in% c('', '\t')]
```

```{r}
base::length(metaphysicsE[! metaphysicsE %in% c('', '\t')])
```

```{r}
stringr::str_length(metaphysicsE)
```

```{r}
base::sum(stringr::str_length(metaphysicsE))
```

```{r}
metaphysicsE_df <- tibble::tibble(section_num = 1:base::length(metaphysicsE), section = metaphysicsE)
```

```{r}
tidy_metaphysicsE <-
  metaphysicsE_df %>%
    tidytext::unnest_tokens(output = word, input = section, to_lower = TRUE) %>%
    dplyr::anti_join(tidytext::stop_words, by = 'word') %>%
    dplyr::anti_join(tibble::tibble(word = base::c('cf', 'ff', 'i.e', 'e.g', '1', '2', '3')), by = 'word')
```
  
```{r}
tidy_metaphysicsE %>%
  dplyr::count(word, sort = TRUE) %>%
  utils::head(10)
```

```{r}
tidy_metaphysicsE %>%
  dplyr::count(word, sort = TRUE) %>%
  dplyr::filter(n > 75) %>%
  dplyr::mutate(word = stats::reorder(word, n)) %>%
  ggplot2::ggplot(ggplot2::aes(n, word)) +
    geom_col() +
    labs(y = NULL)
```

### Token Frequency

```{r}
frequency <- dplyr::bind_rows(dplyr::mutate(tidy_timaeusE,     book = 'Timaeus_Plato'),
                              dplyr::mutate(tidy_physicsE,     book = 'Physics_Aristotle'),
                              dplyr::mutate(tidy_metaphysicsE, book = 'Metaphysics_Aristotle')) %>%
  dplyr::mutate(word = stringr::str_extract(word, "[a-z']+")) %>%
  dplyr::count(book, word) %>%
  dplyr::group_by(book) %>%
  dplyr::mutate(proportion = n / sum(n)) %>%
  dplyr::select(-n) %>%
  tidyr::pivot_wider(names_from = book, values_from = proportion) %>%
  tidyr::pivot_longer(Physics_Aristotle:Metaphysics_Aristotle, names_to = 'Aristotle', values_to = 'proportion')
  
frequency %>%
  head(10)
```

### Correlation

```{r}
stats::cor.test(data = frequency[frequency$Aristotle == 'Metaphysics_Aristotle',], ~ proportion + Timaeus_Plato)
stats::cor.test(data = frequency[frequency$Aristotle == 'Physics_Aristotle',],     ~ proportion + Timaeus_Plato)
```

## Data Visualization

```{r}
ggplot(frequency, aes(x = proportion, y = Timaeus_Plato), color = abs(Timaeus_Plato - proportion)) +
  geom_abline(color = 'gray40', lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), low = 'darkslategray4', high = 'gray75') +
  facet_wrap(~Aristotle, ncol = 2) +
  theme(legend.position = 'none') +
  labs(x = 'Aristotle', y = 'Plato')
```

```{r}
tidy_books <- dplyr::bind_rows(dplyr::mutate(tidy_timaeusE,     book = 'Timaeus_Plato'),
                               dplyr::mutate(tidy_physicsE,     book = 'Physics_Aristotle'),
                               dplyr::mutate(tidy_metaphysicsE, book = 'Metaphysics_Aristotle')) %>%
  dplyr::mutate(word = stringr::str_extract(word, "[a-z']+"))
utils::head(tidy_books)

tidy_books %>%
  dplyr::inner_join(tidytext::get_sentiments('bing')) %>%
  dplyr::count(book, index = section_num, sentiment) %>%
  tidyr::pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  dplyr::mutate(sentiment = positive - negative) %>%
  ggplot2::ggplot(ggplot2::aes(index, sentiment, fill = book)) +
    ggplot2::geom_col(show.legend = FALSE) +
    ggplot2::facet_wrap(~book, ncol = 2, scales = 'free_x')
```

## Conclusion

Answer your research question using the data sets and visualizations you created.